---
title: "Mascoma River Greenway Survey and Count Results"
author: "UVLSRPC"
date: "October 6, 2020"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    theme: yeti
    css: "DashboardStyles.css"
editor_options: 
  chunk_output_type: console
---

```{r ggplot_Theme}
theme_db <- theme_bw() + theme(plot.margin = unit(c(10,10,20,10),"pt"))
```

```{r Map_Settings}
map_type <- "CartoDB.Positron"
```

Overview {data-icon="fa-home"}
============================================

Summary {data-width=200}
--------------------------------------------

### Number of Respondents

```{r}
value <- format(mrg[, .N], big.mark = ",", trim = TRUE)
valueBox(value, caption = "Number of Completed Surveys", icon = "fa-check")
```

### Number of Respondents from Lebanon

```{r}
value <- format(mrg[Muni_State %in% c("WEST LEBANON, NH", "LEBANON, NH"), .N], big.mark = ",", trim = TRUE)
valueBox(value, caption = "Surveys from City of Lebanon Residents", icon = "fa-check")
```

### About this Document

The dashboard presents analyis of the Mascoma River Greenway user survey and counts. The survey was administered to greenway users and other Upper Valley residents in June and July, 2020 by UVLSRPC. Counts at two locations on the Mascoma River Greenway were taken by UVLSRPC for two weeks in late July and early August, 2020.

The survey and count program were developed in collaboration with the Mascoma River Greenway coalition, City of Lebanon, Friends of Lebanon Recreation, Lebanon Pedestrian and Bicyclist Advisory Committee, Upper Valley Trails Alliance, and other groups.

The purpose of the survey and counts were to collect data to understand current use of the greenway, identify any short term issues and user needs, and to support longer term planning of improvements to the greenway.

Users may navigate to different areas of the dashboard using the navigation bar at the top of the page, and may interact directly with most tables, charts, and maps.

This document is best viewed using the most recent versions of the following web browsers:

* [Google Chrome](https://www.google.com/chrome/browser/desktop/)
* [Mozilla Firefox](https://www.mozilla.org/en-US/firefox/new/)
* Microsoft Edge (Must be running Windows 10)


Column {data-width=400}
--------------------------------------------

### Demographics of Respondents

The majority of respondents who completed the survey were from Lebanon, with about two thirds living in the city. Others in the Upper were also well respresented, and there were also almost a 100 responses from further afield in New Hampshire, Vermont and other states.

The proportion of survey respondents by age peaked in the 41 to 60 age bracket. There were relatively few responses from younger MRG users (under 25); the survey asked about the age of the respondent and did not include questions about the age of those accompanying the respondents on their trips on the MRG, so we don't have information about, for example, the number of children using the MRG.

The household income of survey respondents was relatively high on average, with about a third of respondents resporting that they live in households with an annual income of over $100,000. 

### Respondents by Home Location

```{r Chart_Resp_by_Home}
# Organize
dat <- mrg[, .N, by = Muni_State]

# Save
fwrite(dat, file = "Resp_by_Home.csv")

# Plot
bar_plotter(dat, xvar = "Muni_State", yvar = "N", xlabel = "Home Municipality", ylabel = "Respondents", coord_flip = TRUE)
```

### Respondents by Age

```{r Chart_Resp_by_Age}
# Organize
dat <- mrg[,.N, keyby = .(Age = Q_StandardAge)]

# Save
fwrite(dat, file = "Resp_by_Age.csv")

# Plot
bar_plotter(dat, xvar = "Age", yvar = "N", ylabel = "Respondents", coord_flip = TRUE)
```

### Respondents by Household Income

```{r Chart_Resp_by_Income}
# Organize
dat <- mrg[, .N, keyby = .(Income = XIT_Custom3)]

# Save
fwrite(dat, file = "Resp_by_Income.csv")

# Plot
bar_plotter(dat, xvar = "Income", yvar = "N", xlabel = "Household Income", ylabel = "Respondents", coord_flip = TRUE)
```

Column {data-width=400}
--------------------------------------------

### Types of Use and Frequency

Most respondents reported using the MRG for recreation, with lower intensity recreation (walking and leisurely paced cycling) the most common uses. More intensive recreation on foot (running) and bikes was also commonly reported. Smaller number of users reported winter sports uses and alternative warmers months users such as rollerblading. 

About two thirds of respondents only ever use MRG recreation. Amongst non-recreational uses, running errands was the most common response, closely followed by commuting to work or school.

Respondents were asked about their frequency of use in both the warmer months of the year and in winter. Amongst the survey respondents, about half used the MRG either weekly or a few times a week. Winter use is much less frequent, with over a third of respondents never using the MRG in winter and the most common response being "about once a month".


### Recreational Use

Which of the following recreational activities do you use the Greenway for? Check all that apply.

```{r Chart_Rec_Use}
# Organize
dat <- mrg_rec_use[,.(Yes = sum(Yes, na.rm = TRUE)), keyby = .(ResponseOption, UserType)]

# Save
fwrite(dat, file = "Rec_Use.csv")

# Plot
bar_plotter(dat, xvar = "ResponseOption", yvar = "Yes", xlabel = "Recreational Use", ylabel = "Respondents", fill = "UserType", coord_flip = TRUE)
```

### Non-Recreational Use

Which of the following non-recreational activities do you use the Greenway for? Check all that apply.

```{r Chart_Trans_Use}
# Organize
dat <- mrg_nonrec_use[,.(Yes = sum(Yes, na.rm = TRUE)), keyby = .(ResponseOption, UserType)]

# Save
fwrite(dat, file = "Trans_Use.csv")

# Plot
bar_plotter(dat, xvar = "ResponseOption", yvar = "Yes", xlabel = "Non-Recreatonal Use", ylabel = "Respondents", fill = "UserType", coord_flip = TRUE)
```

### Frequency by Season

In the warmer months, about how often do you use the Greenway?

In the winter, about how often do you use the Greenway?

```{r Chart_Freq}
# Organize
dat <- rbind(mrg[,.N, keyby = .(Frequency = S2_P1_T0_Q1_Seasonal_Use_1)][, Season := "Warmer Months"],
             mrg[,.N, keyby = .(Frequency = S2_P1_T0_Q2_Seasonal_Use_2)][, Season := "Winter"])
# Save
fwrite(dat, file = "Freq_Warmer_Months_And_Winter.csv")

# Plot
bar_plotter(dat, xvar = "Frequency", yvar = "N", ylabel = "Respondents", fill = "Season", position = "dodge", xrotate = TRUE)
```

Experience {data-icon="fa-user"}
============================================

Column {data-width=333}
--------------------------------------------

### Experience

Most respondents reported postive experiences when using the MRG, with the most common negative experience being due to a lack of signage and rules. A small number of respondents reported collisions or near misses with bicyclists or other conflicts. The most common request made in the written responses to the specific experience question was to add signs with safety instructions and rules. The most common conflicts or issues were with bicyclists and dogs.

### Overall Experience

Please select the options below that describe your experience using the Greenway

```{r Chart_Experience}
# Organize
dat <- mrg_experience[,.(Yes = sum(Yes, na.rm = TRUE)), keyby = .(ExperienceType, ResponseOptionShort)][order(-Yes)]
dat[, ResponseOptionShort := factor(ResponseOptionShort, levels = dat$ResponseOptionShort)]

# Save
fwrite(dat, file = "MRG_Experience.csv")

# Plot
bar_plotter(dat, xvar = "ResponseOptionShort", yvar = "Yes", fill = "ExperienceType", xlabel = "Experience", ylabel = "Respondents", coord_flip = TRUE)
```

### Specific Experiences

Please describe any specific experiences or thoughts about the Greenway that you wish to share

```{r Chart_Experience_Specific}
# Organize
dat <- mrg_exp_qual[!is.na(Code),.(Yes = .N), keyby = CodedResponse][order(-Yes)]
dat[, CodedResponse := factor(CodedResponse, levels = dat$CodedResponse)]

# Save
fwrite(dat, file = "MRG_Experience_Qual.csv")

# Plot
bar_plotter(dat, xvar = "CodedResponse", yvar = "Yes", xlabel = "Experience or Suggestion", ylabel = "Respondents", coord_flip = TRUE)
```


Column {data-width=333}
--------------------------------------------

### Access to the MRG

Most respondents drove in their car to the MRG, but amongst Lebanon residents, there was more non-motorized access (walking or biking) than driving.

Respondents reported distance as the main barrier to walking or driving to the MRG, followed by unsafe roads. 

### Travel to the MRG

How do you usually travel to the Greenway from your home?

```{r Chart_Travel_Mode}
# Organize
dat <- mrg[,.N, keyby = .(Mode = S2_P3_T0_Q1_Getting_to_the_Greenway_1, LebRes)]

# Save
fwrite(dat, file = "MRG_Travel_Mode.csv")

# Plot
bar_plotter(dat, xvar = "Mode", yvar = "N", ylabel = "Respondents", fill = "LebRes", position = "dodge", xrotate = TRUE)
```

### Barriers to Walk/Bike Access

What are the top TWO barriers that might discourage you from walking or bicycling to the Greenway from your home?

```{r Chart_Barriers}
# Organize
dat <- mrg_barriers[,.(Yes = sum(Yes, na.rm = TRUE)), keyby = .(ResponseOptionShort, BarrierType)]

# Save
fwrite(dat, file = "MRG_Barriers.csv")

# Plot
bar_plotter(dat, xvar = "ResponseOptionShort", yvar = "Yes", fill = "BarrierType", xlabel = "Barrier", ylabel = "Respondents", coord_flip = TRUE)
```

Column {data-width=333}
--------------------------------------------

### Destinations

While many respondents report staying on the MRG, a significant number used the MRG to access destinations immediately to the east including downtown Lebanon and the Northern Rail Trail. The Miracle Mile was also an important destination, followed by West Lebanon (either downtown or 12A).

The most frequently reported barrier to reaching destinations was the lack of a connection to West Lebanon and the poor safety of Glen Road, the Miracle Mile, and the intersection at High Street and Mascoma Street.

### Destinations Visited

Do you currently use the Greenway to help you get to any of these destinations by foot or bike (check ALL that apply)?

```{r Chart_Destinations}
# Organize
dat <- mrg_dest[,.(Yes = sum(Yes, na.rm = TRUE)), keyby = .(ResponseOption, DestType)]

# Save
fwrite(dat, file = "MRG_Destinations.csv")

# Plot
bar_plotter(dat, xvar = "ResponseOption", yvar = "Yes", fill = "DestType", xlabel = "Destination", ylabel = "Respondents", coord_flip = TRUE)
```

### Destination Barriers

Are there any barriers or safety issues that discourage you from using the Greenway to get to local destinations by foot or bike? Please describe below.

```{r Chart_Destination_Barriers}
# Organize
dat <- mrg_dest_qual[!is.na(Code),.(Yes = .N), keyby = CodedResponse][order(-Yes)]
dat[, CodedResponse := factor(CodedResponse, levels = dat$CodedResponse)]

# Save
fwrite(dat, file = "MRG_Destination_Barriers.csv")

# Plot
bar_plotter(dat, xvar = "CodedResponse", yvar = "Yes", xlabel = "Barrier", ylabel = "Respondents", coord_flip = TRUE)
```

Access Maps {data-icon="fa-map"}
============================================

Column {data-width=300}
--------------------------------------------

### Current Access, Desired Access, and Access Barriers

Respondents added maps pins and comments noting the current MRG access locations that they use, their desired access locations, and access barriers that they face.

The most commonly used current access locations are the public access points at each end of the MRG (the east entrance close to downtown Lebanon and at the parking lots at Slayton Hill and APD, and at the western end of the MRG at Glen Road). The access from the Price Chopper parking lot is also used frequently and an agreement with the landowner to formalize this access point is under negotiation by the City of Lebanon.

### Current Access Locations

```{r Chart_Current_Access}
# Organize
dat <- mrg_access
dat[, name := factor(name, levels = dat$name)]

# Save
fwrite(dat, file = "MRG_Current_Access.csv")

# Plot
bar_plotter(dat, xvar = "name", yvar = "Respondents", fill = "AccessType", xlabel = "Acccess", ylabel = "Respondents", coord_flip = TRUE)
```

### Desired Access Locations

```{r Chart_Desired_Access}
# Organize
dat <- mrg_desired[!is.na(Respondents)]
dat[, name := factor(name, levels = dat$name)]

# Save
fwrite(dat, file = "MRG_Desired_Access.csv")

# Plot
bar_plotter(dat, xvar = "name", yvar = "Respondents", xlabel = "Acccess", ylabel = "Respondents", coord_flip = TRUE)
```

Column {data-width=700}
--------------------------------------------

### Access Map

```{r Map_Access}

tmap_mode("view")

tm_basemap("OpenStreetMap") +
tm_shape(access_shp) +
 tm_bubbles(size = "Respondents" ,
           col = "AccessType",
           scale = 10,
           style = "cont",
           id = "name",
           title.col = "Current Access Type",
           palette = rgb(mrgcolordf[2:4,],maxColorValue = 255)) +
tm_shape(desired_barriers_shp) +
 tm_bubbles(size = "Respondents" ,
           col = "DesiredOrBarrier",
           scale = 5,
           style = "cont",
           id = "name",
           title.col = "Desired Access/Access Barrier",
           popup.vars = "Respondents",
           palette = rgb(mrgcolordf[c(1,5),],maxColorValue = 255))
    
  
```


MRG Value {data-icon="fa-dollar-sign"}
============================================

Column {data-width=400}
--------------------------------------------

### The MRG as a Community Resource

Respondents to the survey most strongly value the MRG as a safe, car free location to walk and bike. They also strongly value it because it helps them to get outdoors and enjoy the scenery in a healthy way. The MRG also provides a bike and pedestrian connection through Lebanon.

Value is placed on it being flat, paved and smooth, and that is provides a recreation location for residents with limited mobility.

### Community Resource Values

Do you think the Mascoma River Greenway is a valuable community resource? Why or why not? Please explain below

```{r Chart_Resource}
# Organize
dat <- mrg_comm_val[!is.na(Code),.(Yes = .N), keyby = CodedResponse][order(-Yes)]
dat[, CodedResponse := factor(CodedResponse, levels = dat$CodedResponse)]

# Save
fwrite(dat, file = "MRG_Resource.csv")

# Plot
bar_plotter(dat, xvar = "CodedResponse", yvar = "Yes", xlabel = "Value", ylabel = "Respondents", coord_flip = TRUE)
```

Column {data-width=400}
--------------------------------------------

### MRG Investment Priorities

Respondents to the survey were asked to allocate a notional budget to various MRG maintenance and improvement options. The largest proportion of the budget was allocated to extending the MRG to West Lebanon, recognizing both the preference amongst respodents for an extension and also the likely cost associated with that project. A close second was a shorter extension from the current western terminus of the MRG to Riverside Park.

### Budget

```{r Chart_Budget}
# Organize
dat <- mrg_budget[,.(Amount = sum(Amount, na.rm = TRUE)), keyby = ResponseOption]

# Save
fwrite(dat, file = "MRG_Budget.csv")

# Plot
bar_plotter(dat, xvar = "ResponseOption", yvar = "Amount", xlabel = "Budget Item", coord_flip = TRUE)
```

Counts {data-icon="fa-line-chart"}
============================================

Summary {data-width=200}
--------------------------------------------

### Average Daily Users (East of I-89)

```{r}
value <- format(round(mean(count_mrg[Location == "East of I-89",.(Count = round(sum(UsersEst))), 
          keyby = .(Date)]$Count)), big.mark = ",", trim = TRUE)
valueBox(value, caption = "Daily users (East of I-89) ", icon = "fa-check")
```

### Average Daily Users (Price Chopper)

```{r}
value <- format(round(mean(count_mrg[Location == "Price Chopper",.(Count = round(sum(UsersEst))), 
          keyby = .(Date)]$Count)), big.mark = ",", trim = TRUE)
valueBox(value, caption = "Daily users (Price Chopper) ", icon = "fa-check")
```

### Proportion Pedestrians

```{r}
value <- format("60%", trim = TRUE)
valueBox(value, caption = "Proportion of users that are pedestrians", icon = "fa-walking")
```

### Proportion Cycling

```{r}
value <- format("34%", trim = TRUE)
valueBox(value, caption = "Proportion of users that are cycling", icon = "fa-walking")
```

### Busiest Hour

```{r}
value <- format(max(count_mrg[,.(Count = round(sum(UsersEst)/.N)), 
          keyby = .(Hour,Date,Location)]$Count), trim = TRUE)
valueBox(value, caption = "Highest hourly count", icon = "fa-check")
```

### MRG Count Methodology

UVLSRPC recorded counts at two locations on the Mascoma River Greenway in late July and early August, 2020. The count period lasted for 19 days between July 25th and August 12th and included 6 weekend days and 13 weekdays. The two locations were "East of I-89", between the MRG parking lot at Slayton Hill and the bridge over I-89, and just to the west of the access from the "Price Chopper" parking lot on the Miracle Mile.

The counts are presented in terms of number of unique MRG users. Since most users are making an out and back trip of some kind the counts were factored down to remove duplicate counts of the same people. UVLSRPC also carried out a 3-hour manual calibration count to allow for adjustments of the automated counters, which can under count groups of users if people are walking or biking very close together. The results of the calibration count were used to factor up the counts.

The calibration count also recorded a split between walkers/runners and cyclists of 60% on foot, 34% cycling, with 6% other (all rollerbladers except for one municipal employee on an ATV)

Column {data-width=400}
--------------------------------------------

### Daily Counts

The daily charts below show that, on an average day, about 240 people use the section of the MRG east of I-89 and about 150 people use the section close to the Price Chopper access. The number of daily users does not vary between weekdays and weekends. The section east of I-89 may be busier as it closer to the more densely populated residential areas around downtown Lebanon and there are more formal access points and trailhead parking lots. 

### Daily Count: East of I-89

```{r Chart_Count_Daily_East_I89}
# Organize
dat <- count_mrg[Location == "East of I-89",.(Count = round(sum(UsersEst))), 
             keyby = .(Date, Weekend)]
dat[, DateFac := factor(Date)]

# Save
fwrite(dat, file = "MRG_Count_Daily_East_of_I89.csv")

# Plot
bar_plotter(dat, xvar = "DateFac", yvar = "Count", fill = "Weekend", xlabel = "Date", ylabel = "Count", xrotate = TRUE)
```

### Daily Count: Price Chopper

```{r Chart_Count_Daily_PC}
# Organize
dat <- count_mrg[Location == "Price Chopper",.(Count = round(sum(UsersEst))), 
             keyby = .(Date, Weekend)]
dat[, DateFac := factor(Date)]

# Save
fwrite(dat, file = "MRG_Count_Daily_Price_Chopper.csv")

# Plot
bar_plotter(dat, xvar = "DateFac", yvar = "Count", fill = "Weekend", xlabel = "Date", ylabel = "Count", xrotate = TRUE)
```

Column {data-width=400}
--------------------------------------------

### Impact of High Temperatures and Rain

While the count period had generally consistently dry and warm summer weather, there were some weather events that reduced trail use. On Tuesday August 4th, Tropical Storm Isaias passed through NH and Lebanon saw 1.4 inches of rain, and that day had only around 30 percent of the typical number of users. There was measurable rain (a tenth of an inch or more) on three other days (the four days marked in blue on the chart below were wet). Two days saw excessive heat, with high temperatures of 93F and 94F respectively (these two days are marked in red on the chart below). The second chart below shows that there is some correlation between the daily high temperature and the amount of trail use, with a decline in use as the temperature increases. The range of daily high temperatures in the count period was 78F to 94F.

### Daily Weather

```{r Chart_Count_Daily_Weather}
# Organize
dat <- count_mrg[,.(Count = round(sum(UsersEst)), Temperature = max(Temperature), Precip = sum(Precip)/2), 
             keyby = .(Date)]
dat[, DateFac := factor(Date)]
dat[, Hot := ifelse(Temperature > 92,1,0)]
dat[, Wet := ifelse(Precip > 0,1,0)]
dat[, Weather := ifelse(Hot == 1 & Wet == 1, "Hot and Rain", ifelse(Hot == 1, "Hot and Dry", ifelse(Wet == 1, "Rain", "Dry and not Hot")))]
dat[, Weather := factor(Weather, levels = c("Rain", "Dry and not Hot", "Hot and Dry"))]

# Save
fwrite(dat, file = "MRG_Count_Daily_Weather.csv")

# Plot
bar_plotter(dat, xvar = "DateFac", yvar = "Count", fill = "Weather", xlabel = "Date", ylabel = "Count", xrotate = TRUE)

```

### Daily Max Temperature

```{r Chart_Count_Daily_Max_Temp}
# Organize
dat <- count_mrg[,.(Count = round(sum(UsersEst)), Temperature = max(Temperature), Precip = sum(Precip)), 
             keyby = .(Date, Location)]

# Save
fwrite(dat, file = "MRG_Count_Daily_Max_Temp.csv")

# Plot
p <- ggplot(data=dat[Precip == 0], aes(y = Count, x = Temperature, color = Location)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylab("Daily Count") +
  xlab("Maximum Temperature (F)") +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_x_continuous(labels = scales::comma_format()) +
  scale_color_manual(values = rgb(mrgcolordf[2:3,],maxColorValue = 255)) +
  theme_db
ggplotly(p)


```

Column {data-width=500}
--------------------------------------------

### Hourly Profiles

The two charts below show the average hourly counts at the two locations, split between weekdays in blue and weekend days in green. While the daily use is consistent between weekdays and weekends, the times of day that see peak trail use vary between weekdays and weekends. 

Trail use builds quickly early on weekday mornings, with high levels of use in the early morning hours from 6am to 8am and then contuining at that level until 11am. Lunchtime and early afternoon sees a decline in use, with a second higher peak in the late afternoon and early evening from 4pm until 7pm.

Things take a bit longer to get going on weekends with the 5am to 7am hours being much quieter than weekdays. Later in the morning is similar to weekdays and then there is a peak at 11am and a second bump in use after lunch, from 1pm to 3pm. Evening use is slightly lower than the afternoon and lower than evening use on weekdays.

### Hourly Count: East of I-89

```{r Chart_Count_Hourly_East_I89}
# Organize
dat <- count_mrg[Location == "East of I-89",.(Count = round(sum(UsersEst)/.N)), 
             keyby = .(Hour, Weekend)]
dat[, HourFac := factor(Hour)]

# Save
fwrite(dat, file = "MRG_Count_Hourly_East_of_I89.csv")

# Plot
bar_plotter(dat, xvar = "HourFac", yvar = "Count", fill = "Weekend", position = "dodge", xlabel = "Hour", ylabel = "Count")
```

### Hourly Count: Price Chopper

```{r Chart_Count_Hourly_PC}
# Organize
dat <- count_mrg[Location == "Price Chopper",.(Count = round(sum(UsersEst)/.N)), 
             keyby = .(Hour, Weekend)]
dat[, HourFac := factor(Hour)]

# Save
fwrite(dat, file = "MRG_Count_Hourly_Price_Chopper.csv")

# Plot
bar_plotter(dat, xvar = "HourFac", yvar = "Count", fill = "Weekend", position = "dodge", xlabel = "Hour", ylabel = "Count")
```
